Implementing Capon Adaptive Beamforming on the GPU for Real Time Cardiac Ultrasound Imaging

### 2012-05-24 HRUIonGPU/jrnl_article.tex
article writing jp

\subsection{GPU Compute Model}
In this section we briefly discuss the principle behind GPU computing. CUDA by Nvidia has been our choice of GPU developer environment \cite{Nvidia2011}. However, even though CUDA has been used to implement the Capon beamformer, the challenges and proposed methods described in this paper applies for other massive parallel implementation environments like e.g. OpenCL as well. The CUDA compute model is based on execution of a kernel-function across a grid of compute threads. Each position in the grid holds a block of threads.
%Each block is further divided into groups of 32 threads, known as a warp. The maximum number of threads inside a block is restricted to 512, hence 16 warps. 
From a CUDA perspective the GPU consist of one or more stream multi processors (SM's), where each SM is capable of 32 or more, depending on the architecture, concurrent multiplications and/or additions. A group of blocks is scheduled to each SM until all blocks have been processed. More important, the SM has a limited amount of near-core memory and registers available. Each thread therefore has to be as light on memory as possible and has to perform enough instructions per used byte in order to hide memory latency. This pose a real challenge as the target problem has to map to the parallel framework both on a grid, block and memory consumption level in order to benefits from parallel acceleration.

@@

\section{Parallel Capon}\label{sec:meth}
As mentioned, the widespread use of Capon adaptive beamforming has been limited by its computational complexity. However, the calculation of Capon weights is well suited for a GPU implementation, since the calculations are both independent across pixels, and partially intra-pixel. On a top level it is trivial to recognise that one grid cell of threads should process one pixel in the final image. Further, on a pixel-level, the algorithm can be broken down into three main steps. For each data vector $\vec{x}$ in the image:
\begin{enumerate}
\item Build the sample covariance matrix $\mat{\hat{R}}$
\item Solve the linear system of equations $\mat{\hat{R}}\vec{b} =\vec{ 1}$
\item Calculate $\vec{w} = \vec{b}/|\vec{b}|$ and the amplitude $\vec{w}^H\vec{x}$
\end{enumerate}
The computational complexity of these three steps, with $L = M/2$, is $O(L^3)$, $O(L^3)$, and $O(L)$ respectively. For larger $L$'s, the first two steps quickly becomes unmanageable, and to achieve real time performance it is therefore important to find novel solutions to these steps.

>>>>

\subsection{GPU Compute Model}
In this section we briefly discuss the principle behind GPU computing. CUDA by Nvidia has been our choice of GPU developer environment \cite{Nvidia2011}. However, even though CUDA has been used to implement the Capon beamformer, the challenges and proposed methods described in this paper applies for other massive parallel implementation environments like e.g. OpenCL as well.

The CUDA compute model is based on execution of a kernel-function across a grid of compute threads. Each position in the grid holds a block of threads.
%Each block is further divided into groups of 32 threads, known as a warp. The maximum number of threads inside a block is restricted to 512, hence 16 warps. 
From a CUDA perspective the GPU consist of one or more stream multi processors (SM's), where each SM is capable of 32 or more, depending on the architecture, concurrent multiplications and/or additions. A group of blocks is scheduled to each SM until all blocks have been processed. More important, the SM has a limited amount of near-core memory and registers available. Each thread therefore has to be as light on memory as possible and has to perform enough instructions per used byte in order to hide memory latency. This pose a real challenge as the target problem has to map to the parallel framework both on a grid, block and memory consumption level in order to benefits from parallel acceleration.

@@

\section{Parallel Capon}\label{sec:meth}
As mentioned, widespread use of Capon adaptive beamforming has been limited by its computational complexity. However, the calculation of Capon weights is well suited for a GPU implementation, since the calculations are both independent across pixels, and partially intra-pixel. On a top level it is trivial to recognise that one grid cell of threads should process one pixel in the final image. Continuing, on a pixel-level, the algorithm can be broken down into three main steps that to some extent can be further divided into a small set of parallel processes. For each data vector $\vec{x}$ in the image:
\begin{enumerate}
\item Build the sample covariance matrix $\mat{\hat{R}}$
\item Solve the linear system of equations $\mat{\hat{R}}\vec{b} =\vec{ 1}$
\item Calculate $\vec{w} = \vec{b}/|\vec{b}|$ and the amplitude $\vec{w}^H\vec{x}$
\end{enumerate}
The computational complexity of these three steps, with $L = M/2$, is $O(L^3)$, $O(L^3)$, and $O(L)$ respectively. For larger $L$'s, the first two steps quickly becomes unmanageable, and to achieve real time performance it is therefore important to find novel solutions to these steps.
